Inferring true changes in microbial abundance from taxonomically-biased microbiome measurements

### 2021-12-02 appendix.Rmd
Rename implications section

The final number of reads also varies due to idiosyncratic events that we typically think of (and model) as 'random', such as the precise handling of a sample during pipetting or loading of a sequencing flow cell.
Such variation may or may not be dwarfed by the 'random' variation in the sample composition (i.e., the variation not explained by the covariates in a regression analysis).
For simplicity, we ignore this random variation when considering the impact of taxonomic bias on MGS measurement.
The effect of random variation in regression analysis is addressed in Appendix
\@ref(appendix-regression) and the effect of random counting error is discussed in Section \@ref(implications).

>>>>

The final number of reads also varies due to idiosyncratic events that we typically think of (and model) as 'random', such as the precise handling of a sample during pipetting or loading of a sequencing flow cell.
Such variation may or may not be dwarfed by the 'random' variation in the sample composition (i.e., the variation not explained by the covariates in a regression analysis).
For simplicity, we ignore this random variation when considering the impact of taxonomic bias on MGS measurement.
The effect of random variation in regression analysis is addressed in Appendix \@ref(appendix-regression) and the effect of random counting error is discussed in Section \@ref(case-studies).

### 2021-12-02 solutions.Rmd
Rename implications section

The calibrated compositions can be used for arbitrary downstream microbiome analyses, including relative and absolute DA analysis.
To demonstrate the potential for calibration to improve fold-change measurements, we estimated bias from one sample in the mock community data from @brooks2015thet that contained all 7 species and used it to calibrate the compositions of all samples (Figure \@ref(fig:calibration-example)).
We then estimated species' densities using the uncalibrated and calibrated proportions using the community normalization method, treating the total density as known and fixed.
Visual inspection shows that the bias estimated from just a single control community with all species can be sufficient to greatly reduce the effect of taxonomic bias on the measured fold change in species' proportions and densities (Figure \@ref(fig:calibration-example)).
To demonstrate a practical application, we used the bias estimated from the full set of mock communities as the basis for calibrating the vaginal community time series measurements from the MOMS-PI study (@fettweis2019thev) that were made using the same 16S sequencing protocol.
We imputed the efficiencies of other species using taxonomic relationships (Methods).
Calibration then allowed us to explore the impact of bias on the measured trajectories of species proportions within women over the course of pregnancy as described in Section \@ref(implications).

>>>>

The calibrated compositions can be used for arbitrary downstream microbiome analyses, including relative and absolute DA analysis.
To demonstrate the potential for calibration to improve fold-change measurements, we estimated bias from one sample in the mock community data from @brooks2015thet that contained all 7 species and used it to calibrate the compositions of all samples (Figure \@ref(fig:calibration-example)).
We then estimated species' densities using the uncalibrated and calibrated proportions using the community normalization method, treating the total density as known and fixed.
Visual inspection shows that the bias estimated from just a single control community with all species can be sufficient to greatly reduce the effect of taxonomic bias on the measured fold change in species' proportions and densities (Figure \@ref(fig:calibration-example)).
To demonstrate a practical application, we used the bias estimated from the full set of mock communities as the basis for calibrating the vaginal community time series measurements from the MOMS-PI study (@fettweis2019thev) that were made using the same 16S sequencing protocol.
We imputed the efficiencies of other species using taxonomic relationships (Methods).
Calibration then allowed us to explore the impact of bias on the measured trajectories of species proportions within women over the course of pregnancy as described in Section \@ref(case-studies).
